Supplemental Material for: The evolution of gene expression in the term placenta of viviparous mammals

### 2016-09-28 manuscript/supplemental_material.Rnw
add pathway plots for EGF and actin cytoskeleton

<<placenta_class_table,echo=FALSE,results="hide">>=
placenta_class_table <- function(class="shape") {
    subtable <- placenta.classification.p[,c("rn",class,paste0(class,".fdr")),with=FALSE]
    setnames(subtable,c("Gene","p","FDR"))
    subtable <- head(subtable[FDR <= 0.05,][order(p),],n=50)
    print(xtable(subtable,
                 align=c("c","c","r","r"),
                 display=c("s","s","g","g")
                 ),
          booktabs=TRUE,
          floating=FALSE,
          include.rownames=FALSE,
          math.style.exponents=TRUE,
          size="\\tiny")
}

>>>>

<<placenta_pathway_egid,echo=FALSE,results="hide">>=
load("../data/placenta_core_transcriptome_complete")
plac_path_egid <- function(column) {
    expr <- placenta.core.transcriptome.complete[!is.na(egid),eval(column)]
    names(expr) <- placenta.core.transcriptome.complete[!is.na(egid),egid]
    expr
}

plac_path_genes <- function(genes,column) {
    result <- as.numeric(NULL)
    result[!is.na(genes)] <-
        placenta.core.transcriptome.complete[na.omit(genes),eval(column)]
    names(result) <- names(genes)
    result
}

<<plot_pathway,echo=FALSE,result="hide">>=
library("KEGGgraph")
library("plotrix")
makeAttr <- function(graph, default, valNodeList) {
  tmp <- nodes(graph)
  x <- rep(default, length(tmp)); names(x) <- tmp

  if(!missing(valNodeList)) {
    stopifnot(is.list(valNodeList))
    allnodes <- unlist(valNodeList)
    stopifnot(all(allnodes %in% tmp))
    for(i in seq(valNodeList)) {
      x[valNodeList[[i]]] <- names(valNodeList)[i]
    }
  }
  return(x)
}
graph_attrs <- function(graph,nodecolorcolumn,cutpoint=1) {
    outs <- sapply(edges(graph), length) > 0
    ins <- sapply(inEdges(graph), length) > 0
    ios <- outs | ins
    ## translate the KEGG IDs into Gene Symbol
    ioGeneID <- translateKEGGID2GeneID(names(ios))
    nodesNames <- sapply(mget(ioGeneID, org.Hs.egSYMBOL, ifnotfound=NA), "[[",1)
    names(nodesNames) <- names(ios)
    nodevalue <- plac_path_genes(nodesNames,nodecolorcolumn)
    nodevalue.max <- max(pretty(nodevalue))
    nodevalue.na <- is.na(nodevalue)
    nodevalue[is.na(nodevalue)] <- 0
    nodecolors <- apply(colorRamp(c("gray","blue"))(nodevalue/nodevalue.max),
                        1,
                        function(x){rgb(x[1],x[2],x[3],maxColorValue=255)})
    nodecolors[nodevalue.na] <- "#FFFFFF"
    names(nodecolors) <- names(nodevalue)
    subgraph <- graph #subGraph(names(nodecolors)[nodevalue >= 0.1],
                       #  graph)
    for (node in names(nodevalue)[nodevalue < cutpoint]) {
         subgraph <- removeNode(node,subgraph)
    }
    nAttrs <- list();
    # nAttrs$fillcolor <- makeAttr(graph, "lightgrey", list(orange=names(ios)[ios]))
    nAttrs$label <- nodesNames[nodes(subgraph)]
    nAttrs$fillcolor <- nodecolors[nodes(subgraph)]
    nAttrs$fontsize <- makeAttr(subgraph,14)
    nAttrs$width <- makeAttr(subgraph,2)
    nAttrs$shape <- makeAttr(subgraph,"box")
    list(nAttrs=nAttrs,
         graph=subgraph,
         max=nodevalue.max
         )
}

plot_pathway <- function(pathway="hsa04012",pathcolor=quote(median_expression),cutpoint=1) {

    the.graph <- KEGGgraph::parseKGML2Graph(KEGGgraph::getKGMLurl(pathway))

    graph.attrs <- graph_attrs(the.graph,nodecolorcolumn=pathcolor,cutpoint=cutpoint)
    plot(graph.attrs$graph,"dot",
         nodeAttrs=graph.attrs$nAttrs,
         attrs=list(graph=list(rankdir="LR")))
    color.legend(100,100,200,200,c("0",as.character(graph.attrs$max)),
                 apply(colorRamp(c("gray","blue"))(seq(0,1,length.out=10)),
                       1,
                       function(x){rgb(x[1],x[2],x[3],maxColorValue=255)}))
}

\begin{figure}
<<egf_pathway,echo=FALSE,message=FALSE,fig.width=6,fig.height=6,fig.out.width="\\textwidth">>=
plot_pathway("hsa04012")
@
\caption{EGF KEGG pathway}
\label{fig:egf_kegg_pathway}
\end{figure}

<<actin_pathway,echo=FALSE,results="hide",message=FALSE>>=
library("pathview")
pathview(gene.data=plac_path_egid(quote(as.numeric(median_expression))),
         pathway.id="hsa04810",
         both.dirs=list(gene=FALSE,cpd=FALSE),
         limit=list(gene=200,cpd=29),
         high=list(gene="blue"),
         net.signature=FALSE)
@

\begin{figure}
  \includegraphics{hsa04810.pathview.png}
  \caption{KEGG Actin Cytoskeleton Pathway}
  \label{fig:kegg_actin}
\end{figure}

<<placenta_class_table,echo=FALSE,results="hide">>=
placenta_class_table <- function(class="shape") {
    subtable <- placenta.classification.p[factor==class & type == "aov",list(gene,p,fdr)]
    setnames(subtable,c("Gene","p","FDR"))
    subtable <- head(subtable[FDR <= 0.05,][order(p),],n=50)
    print(xtable(subtable,
                 align=c("c","c","r","r"),
                 display=c("s","s","g","g")
                 ),
          booktabs=TRUE,
          floating=FALSE,
          include.rownames=FALSE,
          math.style.exponents=TRUE,
          size="\\tiny")
}
